---
title: "COVID-19 the case of Qatar"
author: "Panagiotis Stylianos"
date: "04/09/2020"
output: 
  bookdown::html_document2:
    theme: journal
      
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

> This is my first blog-post. In this [repository](https://github.com/etc5523-2020/my-blog-petestylianos) you can find the contents of the blog spot, which can also be found at my [website](https://panagiotis-stylianos.netlify.app)


<span style="font-size:7mm">Required Libraries</span>
```{r pkg-load}
library(coronavirus)
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(gt)
library(DT)
library(captioner)
library(ggthemes)
library(widgetframe)
```

 > In this blog-post I will take a look at the __spread of COVID-19 in Qatar__. This post is part of an assignment for _unit ETC5523 on track for the Master of Business Analytics from Monash University._

<span style="font-size:7mm">Analysis</span>

Let's take a look at our data coming from the [coronavirus package](https://github.com/RamiKrispin/coronavirus). It contains daily data for the number of cases of COVID-19 for a plethora of countries.

First we obtain the relevant data for Qatar.
```{r qatar}
qatar <- coronavirus %>% 
  filter(country == "Qatar") %>% 
  select(c(date, type, cases)) %>% 
  as_tibble()
```

Our dataset contains data for `r nrow(qatar)` days. Starting from _`r qatar$date[1]`_ up to 
_`r qatar$date[nrow(qatar)]`_. The type column contains the following categories: _`r unique(qatar$type)`_.

These are the most recent observations.

```{r tbl- captions, echo = FALSE}
table_nums <- captioner::captioner(prefix = "Tab.")
tab.1_cap <- table_nums(name = "tab_1", 
                        caption = "Daily Cases of COVID-19 in Qatar")
tab.2_cap <- table_nums(name = "tab_2", 
                        caption = "Aggregated Info")
tab.3_cap <- table_nums(name = "tab_3", 
                        caption = "Monthly Data")

```

```{r data-table, fig.cap = tab.1_cap}
tbl1 <- tail(qatar) %>% 
   kable(col.names = c("Date", "Type of Case", "Number of Cases"),label = knitr::opts_chunk$get("label")) %>% 
   kable_styling(font_size = 14, bootstrap_options = "hover", 
                 fixed_thead = list(enabled = T, background = "plum"), position = "center",
                 row_label_position = "c", full_width = T) %>% 
  column_spec(1, width = "10em", italic  = T, border_right = T, monospace = T ) %>% 
  column_spec(2, width = "10em", border_right = T, monospace = T) %>%
  column_spec(3, width = "10em", background = "lightgrey", bold = T, color = "black") %>% 
  footnote(general = "Data from the coronavirus package") 
```

`r table_nums("tab_1")`

```{r table-output, echo = FALSE}
tbl1
```

<span style="font-size:7mm">Let's perform some Exploratory Analysis</span>

```{r figures, fig.cap= 'Cases over time', fig.width= 15, fig.height=8}
p1 <- qatar %>% 
  filter(type %in% c("confirmed", "recovered")) %>% 
  ggplot(aes(date, cases, fill = type)) +
  geom_col() +
  theme_economist(horizontal = TRUE, base_family = "Verdana") +
  scale_fill_economist() +
  scale_color_economist() +
  ggtitle("Cases over time") +
  xlab("Date") +
  ylab("Cases") +
  theme(legend.position = "right", title = element_text(size=12), 
        axis.title.x = element_text(size = 16, angle = 0, face = "bold"),
        axis.title.y = element_text(size = 16, angle = 90, face = "bold"))


p2 <- qatar %>% 
  filter(type %in% c("confirmed", "recovered")) %>% 
  ggplot(aes(date, log(cases), color = type)) +
  geom_line() +theme_economist(horizontal = TRUE, base_family = "Verdana") +
  scale_fill_economist() +
  scale_color_economist() +
  ggtitle("Cases over time in logarithmic scale") +
  xlab("Date") +
  ylab("Cases") +
  theme(legend.position = "right", title = element_text(size=12), 
        axis.title.x = element_text(size = 16, angle = 0, face = "bold"),
        axis.title.y = element_text(size = 16, angle = 90, face = "bold"))

grid.arrange(p1, p2)
```

Let's manipulate our data to get the total cases for each sub-type of case.

```{r stats}
qatar_wide <- qatar %>% 
  pivot_wider(names_from = type, values_from = cases)

stats <- qatar_wide %>% 
  summarise(Total_confirmed = sum(confirmed),
            Total_deaths = sum(death),
            Total_recovered = sum(recovered),
            Max_dead = max(death),
            Max_confirmed = max(confirmed))

```

```{r stats-tbl, fig.cap = tab.2_cap}
st_tbl <- stats %>% 
  kable(col.names = c("Total-Confirmed", "Total-Deaths", "Total-Recovered", "Max-Daily-Deaths", "Max-Daily-Confirmed"), label = knitr::opts_chunk$get("label")) %>% 
   kable_styling(font_size = 14, bootstrap_options = "hover", 
                 fixed_thead = list(enabled = T, background = "plum"), position = "center",
                 row_label_position = "c", full_width = T) %>% 
  # column_spec(1, width = "10em", italic  = T, border_right = T, monospace = T ) %>% 
  # column_spec(2, width = "10em", border_right = T, monospace = T) %>%
  # column_spec(3, width = "10em", background = "lightgrey", bold = T, color = "black") %>% 
  footnote(general = "Aggregated Information") 

```


`r table_nums("tab_2")`

```{r st_tbl, echo=FALSE}
st_tbl
```

Now lets look at the cumulative data.

```{r cumulative-data}
cumul_qatar <- qatar_wide %>% 
  mutate(cuml_confirmed = cumsum(confirmed),
         cuml_death = cumsum(death),
         cuml_recoverd = cumsum(recovered))
```

Using longer format to visualize them.
```{r cumul-vis, fig.cap= 'Cumulative Cases', fig.width= 15, fig.height=8}
cumul_qatar_long <- cumul_qatar %>% 
  pivot_longer(cols = contains("cum"),
               names_to = "type",
               values_to = "cumulative")


cumul_qatar_long %>% 
  ggplot(aes(date, log10(cumulative), color = type)) +
  geom_line() +
  theme_economist(horizontal = TRUE, base_family = "Verdana") +
  scale_fill_economist() +
  scale_color_economist() +
  ggtitle("Cumulative Cases over time") +
  xlab("Date") +
  ylab("Cases") +
  theme(legend.position = "right", title = element_text(size=12), 
        axis.title.x = element_text(size = 16, angle = 0, face = "bold"),
        axis.title.y = element_text(size = 16, angle = 90, face = "bold"))
```

Now, lets explore the disribution of the confirmed cases for each month.

```{r monthly-data}
qatar_monthly <- qatar %>% 
  mutate(month = str_sub(date, 6, 7))


qatar_monthly <-  qatar_monthly %>% 
  mutate(month = recode(month, "01" = "January",
                        "02" = "February",
                        "03" = "March",
                        "04" = "April",
                        "05" = "May",
                        "06" = "June",
                        "07" = "July"))

qatar_monthly$month <- factor(qatar_monthly$month, levels = c("January",
                       "February","March", "April", "May","June","July")) 

```



```{r box-plot, fig.cap= 'Monthly Box-plots', fig.width= 15, fig.height=8}

qatar_monthly %>% 
  filter(type == "confirmed",
         month %in% c("April","May","June","July")) %>% 
  ggplot(aes(month, cases)) +
  geom_boxplot(aes(fill = month)) +
  theme_economist(horizontal = TRUE, base_family = "Verdana") +
  scale_fill_economist() +
  scale_color_economist() +
  ggtitle("Monthly Box-plots") +
  xlab("Month") +
  ylab("Cases") +
  theme(legend.position = "right", title = element_text(size=12), 
        axis.title.x = element_text(size = 16, angle = 0, face = "bold"),
        axis.title.y = element_text(size = 16, angle = 90, face = "bold"))
```


Finally, let's produce an interactive table for users to explore the number of cases based on month and category of case.

```{r interactive-tbl, fig.cap = tab.3_cap}
interactive_tbl <- DT::datatable(
qatar_monthly %>% 
  group_by(month, type) %>% 
  summarise(Total = sum(cases))) 

```

`r table_nums("tab_3")`

```{r show-inter-tbl, echo=FALSE}
widgetframe::frameWidget(interactive_tbl)
```


